Coding, consistency & calibration

### 2024-04-03 paper_01_consistency.Rmd
use native pipe & check whether everything still renders

---
title: "Triangulating iconicity: Coding, consistency & calibration"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","irr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

@@

d1 = read_excel("data/ideophones_coded_MD_original.xlsx") %>% arrange(filename)
d2 = read_excel("data/ideophones_coded_SP_original.xlsx") %>% arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

@@

d.F <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
#write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")

d.M <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
#write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")

# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
get_coding_categories <- function(data) {
  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
  gsub('MD','',coder_1_names)
}

@@

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)

@@

```{r consistency_2ndround}
d1 = read_excel("data/ideophones_coded_MD.xlsx") %>% arrange(filename)
d2 = read_excel("data/ideophones_coded_SP.xlsx") %>% arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

@@

d.F <- d %>%
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) %>%
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end

d.M <- d %>%
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) %>%
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)

@@

inconsistent <- c('F_redup','F_monosyllabic','F_vowelquality','F_finallength')
d.F.inconsistent <- d.F %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) %>%
  mutate(F_redup_C = ifelse(F_redup_MD == F_redup_SP,1,0),
         F_monosyllabic_C = ifelse(F_monosyllabic_MD == F_monosyllabic_SP,1,0),
         F_vowelquality_C = ifelse(F_vowelquality_MD == F_vowelquality_SP,1,0),
         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) %>%
  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) %>%
  filter(check < 4)
d.F.inconsistent <- d.F.inconsistent[, order(names(d.F.inconsistent))]

@@

d.coded <- read_excel("data/ideophones_coded_SP.xlsx") %>% 
  arrange(filename) %>%
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

notes <- d %>% dplyr::select(filename,matches('notes'))

@@

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d.coded %>%
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         )

@@

#sanity check: are they indeed mutually exclusive? these calls should come up empty
d %>% 
  mutate(C_aspect = C_iterative + C_punctual) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) %>%
  filter(C_aspect > 1)
d %>%
  mutate(C_length = C_long + C_closure) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
  filter(C_length > 1)

d <- d %>%
  mutate(C_aspect = C_iterative + C_punctual,
         C_length = C_long + C_closure)

@@

d <- d %>%
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent

# add cumulative iconicity scores (and some others)
# any1 = 1 if an ideophone has at least 1 congruency feature of any type
# atleast2 = if an ideophone has at least 2 congruency features of any type
# ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
d <- d %>%
  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,
                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3

>>>>

---
title: "Coding, consistency & calibration"
author: "[anonymized for review]"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","irr")
# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

@@

d1 = read_excel("data/ideophones_coded_MD_original.xlsx") |> arrange(filename)
d2 = read_excel("data/ideophones_coded_SP_original.xlsx") |> arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

@@

d.F <- d |>
  dplyr::select(c("filename","ideophone"),starts_with("F_")) |>
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
#write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")

d.M <- d |>
  dplyr::select(c("filename","ideophone"),starts_with("M_")) |>
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
#write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")

# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
get_coding_categories <- function(data) {
  coder_1_names <- data |> dplyr::select(ends_with("_MD")) |> names()
  gsub('MD','',coder_1_names)
}

@@

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) |> dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) |> dplyr::rename(category = M_categories)

@@

```{r consistency_2ndround}
d1 = read_excel("data/ideophones_coded_MD.xlsx") |> arrange(filename)
d2 = read_excel("data/ideophones_coded_SP.xlsx") |> arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

@@

d.F <- d |>
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) |>
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end

d.M <- d |>
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) |>
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) |> dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) |> dplyr::rename(category = M_categories)

@@

inconsistent <- c('F_redup','F_monosyllabic','F_vowelquality','F_finallength')
d.F.inconsistent <- d.F |>
  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) |>
  mutate(F_redup_C = ifelse(F_redup_MD == F_redup_SP,1,0),
         F_monosyllabic_C = ifelse(F_monosyllabic_MD == F_monosyllabic_SP,1,0),
         F_vowelquality_C = ifelse(F_vowelquality_MD == F_vowelquality_SP,1,0),
         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) |>
  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) |>
  filter(check < 4)
d.F.inconsistent <- d.F.inconsistent[, order(names(d.F.inconsistent))]

@@

d.coded <- read_excel("data/ideophones_coded_SP.xlsx") |> 
  arrange(filename) |>
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

notes <- d |> dplyr::select(filename,matches('notes'))

@@

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d.coded |>
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         )

@@

#sanity check: are they indeed mutually exclusive? these calls should come up empty
d |> 
  mutate(C_aspect = C_iterative + C_punctual) |>
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) |>
  filter(C_aspect > 1)
d |>
  mutate(C_length = C_long + C_closure) |>
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) |>
  filter(C_length > 1)

d <- d |>
  mutate(C_aspect = C_iterative + C_punctual,
         C_length = C_long + C_closure)

@@

d <- d |>
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent

# add cumulative iconicity scores (and some others)
# any1 = 1 if an ideophone has at least 1 congruency feature of any type
# atleast2 = if an ideophone has at least 2 congruency features of any type
# ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
d <- d |>
  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) |>
  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,
                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3
